---
title: R Basics
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)

# We will learn what this code does in Milestone 2
covid <- read_csv("data/covid.csv")
```

## R Basics

In this milestone, you'll become familiar with the `covid` data set.

## Recreation

### Part 1

The data we will use in our investigation is saved as an R object named `covid`. Use R to inspect the contents of `covid`:

```{r}
#| label: recreation
# Remember that all tidyverse packages have already been installed for you, but you will need to *load* the ones you want to use.
```

-   Display the contents of `covid`

```{r}
glimpse(covid)
```

-   Use `nrow()` to find the number of observations in `covid`

```{r}
nrow(covid)

#!!! 20, 780 observations
```

-   Use `mean()` to find the average number of COVID-19 tests in the data
-   Use `max()` to find the maximum number of COVID-19 deaths in as single day in a single state
-   Use `summary()` to generate a summary of the data set

```{r}
mean(covid$tests)
unique(covid$tests)
#!!! 17, 508 average tests

max(covid$deaths)
#!!! 2,559 maximum deaths/day

summary(covid)
```

*Hint:* you may want to look up the help pages for each of these functions.

## Extension

Explore `covid` in a new way. Consider using a new function, or something that you can learn from a help page. Alternatively, try to learn more about the other variables in `covid`. Keep in mind that we will study how to make plots for the next milestone.

**For your first extension, we encourage you to think small.** Throughout this course, you will build many skills to explore data; for this first week, your work will likely be fairly straightforward. Consider *limiting yourself to 30-45 minutes* for this portion of your work; at the end of that time interval, stop working and make notes on your progress, dead-ends, and aspirations -- bringing those to your group for discussion during the milestone session.

Write your extension code in the following chunk:

```{r}
#| label: extension
#| 

```
